```{r message=FALSE, echo=FALSE}
library(xtable)
options(xtable.type='html')
options(xtable.include.rownames=FALSE)
options(xtable.html.table.attributes='border="1" cellspacing="0"')
library(plyr)
source('../run.R')
source('../plots.R')
source('../RunningAHEAD.R')
load('../data/StumpJump50k_results_2002-2013.Rda')
sjResults <- results
load('../data/UpChuck50k_results_2008-2013.Rda')
ucResults <- results
colCommon <- c('lastname','firstname','time_hour')
z <- merge(subset(ucResults,select=colCommon),subset(sjResults,select=colCommon),by=c('lastname','firstname'))
zNames <- names(z)
zNames[3] <- 'time.uc'
zNames[4] <- 'time.sj'
names(z) <- zNames
data <- ddply(z,.(lastname,firstname),summarize, time.uc.mean=mean(time.uc),nraces.uc=length(unique(time.uc)),time.sj.mean=mean(time.sj),nraces.sj=length(unique(time.sj)))

data$time.uc.recommended <- data$time.sj.mean + (4/6) # 40 minutes

sj.times <- seq(4,9,by=.5)
uc.predicted <- predict(lm(time.uc.mean ~ time.sj.mean,data),data.frame(time.sj.mean=sj.times))
race.prediction <- data.frame(
  StumpJumpTime=sec2str(sj.times*3600),
  DirectorRecommendation=sec2str((sj.times+(4/6))*3600),
  ModelRecommendation=sec2str(uc.predicted*3600)
)



# For larger plots
#
opts_chunk$set(fig.width=8,fig.height=8,message=FALSE,echo=FALSE,warning=FALSE)
myTheme <- theme_grey()
myTheme$plot.title$size <- 18
myTheme$plot.title$face <- "bold"
myTheme$text$size <- 16
myTheme$text$face <- "bold"
myTheme$axis.title.x$size <- 14
myTheme$axis.title.x$vjust <- -.2
myTheme$axis.title.y$size <- 14
myTheme$axis.title.y$vjust <- .2
myTheme$legend.text$size <- 14
myTheme$legend.text$face <- "bold"
myTheme$plot.margin <- unit(rep(1,4),'lines')
theme_set(myTheme)

# For smaller plots
#opts_chunk$set(fig.width=6,fig.height=6,message=FALSE,echo=FALSE,warning=FALSE)
```

Over on 
[Upchuck's Facebook page](https://www.facebook.com/upchuck50k), the organizers commented that a good
rule of thumb for average runners is to add 30-40 minutes to their
StumpJump time.  So, using the finishing times of both races (thanks
again to UltraSignup) I created a very 
[simple prediction model](https://github.com/jeffreyhorner/RUNNING/blob/master/reports/StumpJump_vs_UpChuck_times.Rmd) with a
table and plot below. I found all runners who have finished both races,
averaged their times, and ran them through an OLS model. If you want to
see what a real statistician can do with this data, check out 
[Dr. Chris Fonnesbeck's analysis](http://nbviewer.ipython.org/urls/raw.github.com/fonnesbeck/ultra_race_results/master/Race%20Finishing%20Times.ipynb) for something a little more stringent (much
thanks to him as he's mentoring me). I am showing you the results of my
prediction model below as it's a little less intimidating, and it bears
very nearly the same results as Fonnesbeck's.

I wish I had more time to spend on this, but I've run out and
I wanted to get it out there before the race tomorrow. Head over to the
[Facebook page](https://www.facebook.com/upchuck50k) if you want to add to the discussion.

# Prediction Table

So the table below has two recommendations. I call them
DirectorRecommendation and ModelRecommendation. I've simplified
the Director's recommendation to add 40 minutes as it gives you the
more conservative estimate rather than the 30-40 minute range. So for
example, if you've run the StumpJump in a time of 4 hours, the Director
expects you to finish in 4:40 while the Model expects you to finish in
5:00. Interesting, a difference of 20 minutes. As you continue reading
down the table, you'll see that for a StumpJump time of 5:30, both
the Director and the Model agree that you will finish in about 6:10,
no difference. Reading further if you finished StumpJump in 8:00, the
Director expects 8:40 while the model expects 8:05.


```{r, results='asis'}
print(xtable(race.prediction))
```
<br>
# Plot

The DirectorRecommendation is the red line, while the ModelRecommendation
is the blue line. Start by finding your StumpJump time on the X axis
with your computer mouse, scroll up to either one of the lines, then
scroll to the left to get your predicted UpChuck Time.

```{r}
#stat_smooth(method='lm',formula= y ~ log(x)) +
ggplot(data,aes(x=time.sj.mean,y=time.uc.mean)) + 
geom_point() +
geom_smooth(method=lm,se=FALSE) +
geom_line(data=data,aes(x=time.sj.mean,y=time.uc.recommended),inherit.aes=FALSE,color="red") +
scale_y_continuous(breaks=seq(4,10,by=.5),minor_breaks = seq(3.75,10.75,by=.25)) +
scale_x_continuous(breaks=seq(4,10,by=.5),minor_breaks = seq(3.75,10.75,by=.25)) +
xlab('Average StumpJump Results (hours)') + ylab('Average UpChuck Results (hours)')
```
