```{r message=FALSE, echo=FALSE}
# For larger plots
#
opts_chunk$set(fig.width=10,fig.height=10,message=FALSE,echo=FALSE,warning=FALSE)
myTheme <- theme_grey()
myTheme$plot.title$size <- 18
myTheme$plot.title$face <- "bold"
myTheme$text$size <- 16
myTheme$text$face <- "bold"
myTheme$axis.title.x$size <- 14
myTheme$axis.title.x$vjust <- -.2
myTheme$axis.title.y$size <- 14
myTheme$axis.title.y$vjust <- .2
myTheme$legend.text$size <- 14
myTheme$legend.text$face <- "bold"
myTheme$plot.margin <- unit(rep(1,4),'lines')
myTheme$plot.background$fill <- 'white'
theme_set(myTheme)

# For smaller plots
#opts_chunk$set(fig.width=6,fig.height=6,message=FALSE,echo=FALSE,warning=FALSE)
#
load('../data/WhereTennesseeRuns.RData')
locations <- c('Nashville, TN','Chattanooga, TN','Knoxville, TN','Memphis, TN')
expandBox <- function(b) c(b$lng-.2,b$lat-.15,b$lng+.2,b$lat+.15)
```
# Where Nashville Runs

Over on Flowing Data, Nathan Yau mapped out [Where people
run](http://flowingdata.com/2014/02/05/where-people-run/) in major
cities across the US. I thought I'd do the same with the four major
cities in Tennessee. I used RunKeeper's route search tool to find routes
in Nashville, Chattanooga, Knoxville, and Memphis and then graphed them
altogether on tile maps from CloudMade. See anything interesting.

As Yau pointed out in his post, most people like to run near water or
in state parks. I think this is true here in Tennessee as well, but also
note that many run where they work or live or go to school.

```{r where-nashville-runs, fig.width=10}
createRouteMap(loc=locations[1],routes=subset(df,location==locations[1]))
```

## Nashville (larger map)
```{r where-nashville-runs-large, fig.width=20,fig.height=20}
b <- expandBox(geocode(locations[1],'all')$result[[1]]$geometry$location)
createRouteMap(loc=b,routes=subset(df,location==locations[1]))
```

## Chattanooga
```{r where-chattanoga-runs-large, fig.width=20, fig.height=20}
b <- expandBox(geocode(locations[2],'all')$result[[1]]$geometry$location)
createRouteMap(loc=b,routes=subset(df,location==locations[2]))
```
## Knoxville
```{r where-knoxville-runs-large, fig.width=20, fig.height=20}
b <- expandBox(geocode(locations[3],'all')$result[[1]]$geometry$location)
createRouteMap(loc=b,routes=subset(df,location==locations[3]))
```
## Memphis
```{r where-memphis-runs-large, fig.width=20, fig.height=20}
b <- expandBox(geocode(locations[4],'all')$result[[1]]$geometry$location)
createRouteMap(loc=b,routes=subset(df,location==locations[4]))
```
